Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-04-08 VegOzone.tex
First paragraph in intro edited and change near-surface ozone to ground level ozone

\begin{abstract}
Here we assess the importance of characterization of key meteorological variables and subarctic biomes on the modeled susceptibility of subarctic vegetation to ozone used in pollution and climate risk assessments. Future climate changes will manifest in both direct and indirect effects influencing the species' exposure to air pollution. A likely increase in the frequency of persistent heatwaves in the future will inflict heat-related stress while generating favorable conditions for the formation of near-surface ozone, e.g. due to wildfires.\\
We have monitored near-surface ozone concentrations at a high latitude site in Northern Norway during the growing season 2018/19. Due to data acquisition problems, ozone concentrations for two weeks in July 2018 were missing from our record. We present a reconstruction based on a probability density function (PDF) and concidering the Swedish and Finnish atmospheric monitoring sites. We review available near-surface ozone reanalysis products for the region and find only CAMS regional Ensemble satisfactory.
In terms of anomalies from a multi-annual mean, ozone concentrations were only significantly enhanced in April/May 2018 before the start of the growing season. Temperatures and global irradiance showed consecutively significant positive anomalies during Spring and Summer 2018. In 2019 only early spring and late summer were significantly warmer, while colder during summer. Precipitation was lower in both years but displayed a large variability.\\
We investigate the difference in uptake of ozone using the $\mathrm{DO_3SE}$ model, for typical vegetation at the location. We find that bespoke parameterization for subarctic vegetation is essential for an accurate estimation of Phyto-toxic Ozone Dose (POD) and an assessment of damage risk. Accurate estimates of the start and length of growing season are also crucial (up to $15\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ deviation in $\mathrm{POD_1}$).\\
We find that temperature and its derivative vapor pressure deficit are likely the most important climate variables while the variability of ozone concentration within climatological reasonable limits plays a minor role.
The use of subarctic parameterization suggests an improvement in risk assessment as POD values are increased compared to standard expectations; corroborated by the evidence of visible injury in the Svanhovd ozone garden. We estimate a total biomass reduction in Norway spruce of $2.5-3\,\unit{\%}$, downy birch by $15.5-17.4\,\unit{\%}$ and an above-ground biomass reduction in perennial grassland of $12.4-13.7\,\unit{\%}$.  
\end{abstract}

@@

Ground level ozone is a highly toxic pollutant known to cause damage to a variety of ecosystems around the World \citep{PT:Emberson2020}. At continuously high concentration levels, ozone negatively affects life expectancy of humans \citep{WHO2008} and is a cause of visible injury, photosynthetic damage, early senescence as well as programmed cell death of plants \citep{PCE:Kangasjarvi2005}. Annual global yield losses of four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} and a suggested loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
The crop yield loss studies indicate a threat to food security in rapidly developing countries, e.g., in East and South-East Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018}, where ozone concentrations related to enhanced pollutant emissions from the transport, residential, power generation and industrial sectors have increased in the past few decades. Even though long-term (pre 1950) observations of near-surface ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase until relatively recently \citep[Chapter 2]{IPCC2013}. Recent trend analyses indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric ozone started to level-off or even decline \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}. This can be attributed to a successful implementation of air quality regulations coupled with economic restructuring which saw a move from coal to gas, reducing the number of episodes of peak concentrations especially in summertime over Europe and North America \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}. However, there are indications that an increasing frequency of thermal stress accompanying climate change will reduce vegetation's' capacity for ozone uptake and therefor scotch these achievements \citep{NCC:Lin2020}.\\

@@

For computing the $\mathrm{POD_y}$, an estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf. Details specific to the $\mathrm{DO_3SE}$ model are given in Section~\ref{sec:do3se}. $\mathrm{POD_y}$ is then calculated according to:
%
\begin{equation}
  \mathrm{POD_y} = \int{(\Phi_\mathrm{sto}-y)\cdot \mathrm{dt}},
  \label{eq:pod}
\end{equation}
%
with the hourly averaged near-surface ozone flux through the stomata Eq.~(\ref{eq:flux_stomata}) and a flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}. The flux threshold $\mathrm{y}$ is representing the detoxification potential of the plant and is typically only exceeded during daylight hours (i.e. when global radiation is above $50\,\unit{W\,m^{-2}}$).

@@

Within the scope of this study to assess probable effects of conditions unique to northern Fennoscandia on ozone uptake by plants and resultant enhanced susceptibility to ozone inflicted damage, we conducted near-surface ozone concentration measurements at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in the growing seasons of 2018/19 in collaboration with the Norwegian Institute for Air Research (NILU). 
In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden at Svanhovd. In contrast, no such damage was found on the clovers in 2019, although it was found on the sensitive tobacco cultivar. This indicates that the vegetation may have been more affected by ozone in 2018 than in 2019. In terms of $\mathrm{AOT_{40}}$, the threshold for protection of vegetation of $3000\,\unit{ppb}$ (integrated over the growing season) given by the EU directive has been well exceeded at many ozone monitoring sites in Southern Norway in the 2018 growing season \citep{NILU2019} but not at Svanhovd.

@@

The observation site at NIBIO Environment Centre Svanhovd and all related data are presented in Section~\ref{sec:data}. In Section~\ref{sec:stats}, we derive climatologies for near-surface ozone concentration at Svanhovd and northern Fennoscandia based on in situ observations and also compute climatologies of near-surface temperature, precipitation, and global radiation for Svanhovd. Based on these, we evaluate anomalies of environmental key variables in 2018/19 by statistical means. In addition, the derived near-surface ozone concentration climatology for northern Fennoscandia will be compared with available ozone reanalysis products. We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate main environmental drivers for the observed ozone damage in 2018 (Section~\ref{sec:do3se}). In this context, we introduce the$\mathrm{DO_3SE}$ model equations and derive bespoke parameterizations for subarctic species to assess systematic uncertainties arising from the use of default parameterizations. In Section~\ref{sec:conc} finally, we will summarize our results, draw conclusions, and give an outlook on further work.

@@

\subsection{Other data}
\label{subsec:other_data}
To determine the significance of the ozone conditions in 2018, we derive a near-surface ozone concentration climatology for Svanvik. From 1986--1996 ozone monitoring has been operated by NILU at Svanhovd, but no ozone measurements have been conducted for the past $14\,\unit{years}$. Due to probable changes in northern hemispheric tropospheric background ozone and a reduction of summertime peak values related to ozone precursors which fell under air quality regulations in Europe, we will cross-calibrate our data with long-term ozone observations at other sites in northern Fennoscandia. In regard to ozone, available observational data are sparse in this region. An overview over the geographical locations of conducted atmospheric monitoring in the past and present is given in Fig.~\ref{fig:station_map_fennoscandia}. The data are described in Section~\ref{subsubsec:ebas}. 
As additional complement to observational data which are not available at all times and spatially sparse, we also evaluate various ozone reanalysis products (Section~\ref{subsubsec:ozone_rea}).

@@

\subsubsection{Near-surface ozone observations}
\label{subsubsec:ebas}
All near-surface ozone observations used in this study are available from the EBAS database operated by \citet{NILU_EBAS}. In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking. All stations with available long-term observations are located on higher ground than Svanhovd which will result in higher near-surface ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors \citep[e.g.,][]{AB:Klingberg2009}. The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, and heath. However, at close vicinity it varies substantially. Ozone measurements at Pallas are taken on top of mount Sammaltunturi $100\,\unit{m}$ above tree line (pine forest) and mainly consists of low vascular plants, mosses, and lichen \citep{BER:Hatakka2003}. At Svanhovd, the ozone measurements are taken amidst a meadow surrounded by agriculturally managed herbaceous plants and grass (c.c. Fig.\ref{fig:svanhovd_research_station}).

@@

Details of reanalysis data sets referred to in this work are listed in Table~\ref{tab:ozone_rea}.
There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone (MACC and CAMSRA) \citep{ACP:Inness2013, ACP:Inness2019}. The temporal as well as spatial resolution of these reanalysis products is rather coarse: 3-hourly and $0.75^\circ\times 0.75^\circ$ or roughly $29.3\,\unit{km}\times 83.4\,\unit{km}$ at the location of Svanvik, respectively. In addition, higher resolution near-surface ozone reanalysis for whole Europe is available from Copernicus service for regional air quality control (note: since May 31st 2020 through the CAMS Atmosphere Data Store (ADS)). The regional air quality model reanalysis ensemble (CAMS regional Ensemble) is based on nine European state-of-the-art numerical air quality models \citep{Copernicus_Regio_AIRQ}. The ensemble mean is at fairly high spatial and temporal resolution compared to the global reanalysis from ECMWF: $0.1^\circ\times 0.1^\circ$ or roughly $3.9\,\unit{km}\times 11.1\,\unit{km}$ and 1-hourly, respectively. The covered time periods differ and no data are available for before the turn of the millennium. The CAMSRA and CAMS Regional Air Quality ensemble mean are available in near real time, but only former covers a period of sufficient length for climate analysis (2003 -- present). All reanalyses use different but, at that time, latest versions of the operational weather forecast system (OpenIFS) from ECMWF. They differ substantially in the assimilated observational ozone data. The MACC reanalysis assimilates only satellite derived tropospheric column ozone (TC), while CAMSRA also includes ozone profiles from satellite retrievals. Most importantly for our study, in situ observations from ozone near-surface station networks are only assimilated in the CAMS Regional Ensemble. The quality with respect to observations in northern Fennoscandia will be discussed in Section~\ref{subsubsec:clim_ozone}. Without too much foreshadowing, the results prevented us from fetching the whole CAMSRA data.

@@

Finally, we shall look at the comparison of climatologies derived from observational ozone data and the various reanalysis products (Fig.~\ref{fig:ozone_climatologies}). As representative for the sites, the nearest neighboring grid point has been chosen. The ECMWF products do not reproduce observed near-surface ozone well. The MACC reanalysis (Fig.~\ref{fig:ozone_climatologies}a) is in general too low ($(9\pm 7)\,\unit{ppb}$) and displays an erroneous seasonal cycle. Ozone concentrations peak already in March followed by a second peak in July. Also CAMSRA matches the observed ozone climatology poorly (Fig.~\ref{fig:ozone_climatologies}b). Despite displaying only one ozone peak, it still does not reproduce the seasonality in northern Fennoscandia well. The CAMSRA spring peak appears to lag behind by $1\,\unit{month}$ and is $5\,\unit{ppb}$ lower, whereas the minimum occurs in January compared to August/September. The annual amplitude ($(26\pm 1)\,\unit{ppb}$) is larger than in the climatology derived from observations ($19\,\unit{ppb}$). Modeled \chem{[O_3]} at Svanvik are even slightly higher than at Pallas. In contrast, the ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia quite well (Fig.~\ref{fig:ozone_climatologies}c), but is on average still slightly too low ($(2.8\pm 0.5)\,\unit{ppb}$).

\begin{figure}[t]
  %\includegraphics[width=8.3cm]{ozone_climatologies_portrait}
  \includegraphics[width=8.3cm]{fig07}
  \caption{Comparison of climatologies derived from observational data (EBAS) with various ozone reanalysis products: (a) ECMWF: MACC reanalysis; (b) ECMWF: CAMS reanalysis; (c) Copernicus: Regional model reanalysis ensemble mean. As representative for the sites, the nearest neighboring grid point has been chosen. The ECMWF products do not reproduce observed near-surface ozone well. The MACC reanalysis is in general too low and displays an erroneous seasonal cycle. The CAMS reanalysis matches better but does not reproduce the seasonality in northern Fennoscandia. The ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia, though the modeled ozone concentrations are slightly too low.}
  \label{fig:ozone_climatologies}
\end{figure}

@@

In the following, we will use the interchangeable terms anomalies and residuals to distinguish between a climatological (anomalies) and a statistical (residuals) point of view.
All the above factors make it necessary to look not only at anomalies but statistically assess ozone concentration residuals at Svanvik as well as at Esrange and Pallas in comparison. Since the climatology of latter is more reliable, we will start with near-surface ozone concentration residuals \chem{\Delta[O_3]} at Esrange and Pallas in 2018. As data for 2019 has not yet been available from EBAS (last access, February 2020), we refrain from looking into these.

@@

Because the available reanalysis products for near-surface ozone have their limitations, we deem it necessary to reconstruct the missing data in 2018 for the purpose of $\mathrm{DO_3SE}$ modeling by applying a Reynolds's decomposition of the form:
\begin{equation}
  \chem{[O_3]} = \chem{\left<[O_3]\right>} + \chem{\Delta[O_3]},
  \label{eq:reynolds_decomp}
\end{equation}
with climatology \chem{\left<[O_3]\right>} and anomalies \chem{\Delta[O_3]}.
The exact procedure will be described in the following.
We derive hourly climatologies for individual stations and combine Esrange/Pallas to derive a climatology for northern Fennoscandia for July (Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}a). We compute temporal correlations between Svanvik and the other stations in northern Fennoscandia to estimate a time lag (Supplement Fig.~\ref{fig:time_lag_correlation}. We find $3\,\unit{h}$ for Esrange and Pallas with respect to Svanvik. Due to the high correlation with Svanvik (Appendix~Fig.~\ref{fig:density_distribution}), we decide on Pallas as reference station for the reconstruction. We correct the Fennoscandic climatology for the found time lag \chem{\left<[O_3]\right>_{hourly,\,t-3}} by shifting the respective time series back in time by $3\,\unit{h}$ and use it to derive a projection for Svanvik:
\begin{equation}
  X = \frac{\chem{\left<[O_3]\right>_{hourly}^{Svanvik}}}{\chem{\left<[O_3]\right>_{hourly,\,t-3}}}.
  \label{eq:mapping}
\end{equation}
As an intermediate step, we apply Eq.~(\ref{eq:reynolds_decomp}) to derive hourly anomalies for each station $i$
\begin{equation}
  \chem{\Delta[O_3]_{hourly}}^i = \chem{[O_3]_{hourly}}^i-\chem{\left<[O_3]\right>_{hourly}},
\end{equation}
with $i \in \{\text{Esrange, Pallas, Svanvik\}}$.
In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}b), these hourly ozone concentration anomalies are shown. We also correct the ozone anomalies for Pallas for the time lag \chem{\Delta[O_3]_{hourly,\,t-3}^{Pallas}} and use the mapping function in Eq.~(\ref{eq:mapping}) to derive reconstructed anomalies for the missing values at Svanvik:
\begin{equation}
    \chem{\Delta [O_3]_{hourly}^{Svanvik,\,reco}} = \chem{\Delta[O_3]_{hourly,\,t-3}^{Pallas}} \cdot X
\end{equation}
Finally, we add the bias corrected Svanvik ozone climatology and derive a reconstructed ozone concentration time series \chem{[O_3]_{hourly}^{Svanvik,\,reco}}. In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}c), the result is shown together with the observed data and the Copernicus Regional Model Reanalysis at the nearest grid point for comparison.

@@

We found representative ozone climatologies (multi-annual means) for northern Fennoscandia and Svanhovd based on near-surface ozone monitoring. To account for missing data from our 2018 record, we proposed a routine for gap filling of ozone observations following a Reynolds's decomposition into anomalies and climatology. We also assessed the quality of available global reanalysis (ECMWF MACC and CAM) and regional model reanalysis (Copernicus) near-surface ozone concentrations products and conclude that only the Copernicus product can sufficiently reproduce observed ozone concentrations in northern Fennoscandia.

>>>>

\begin{abstract}
Here we assess the importance of characterization of key meteorological variables and subarctic biomes on the modeled susceptibility of subarctic vegetation to ozone used in pollution and climate risk assessments. Future climate changes will manifest in both direct and indirect effects influencing the species' exposure to air pollution. A likely increase in the frequency of persistent heatwaves in the future will inflict heat-related stress while generating favorable conditions for the formation of ground level ozone, e.g. due to wildfires.\\
We have monitored ground level ozone concentrations at a high latitude site in Northern Norway during the growing season 2018/19. Due to data acquisition problems, ozone concentrations for two weeks in July 2018 were missing from our record. We present a reconstruction based on a probability density function (PDF) and concidering the Swedish and Finnish atmospheric monitoring sites. We review available ground level ozone reanalysis products for the region and find only CAMS regional Ensemble satisfactory.
In terms of anomalies from a multi-annual mean, ozone concentrations were only significantly enhanced in April/May 2018 before the start of the growing season. Temperatures and global irradiance showed consecutively significant positive anomalies during Spring and Summer 2018. In 2019 only early spring and late summer were significantly warmer, while colder during summer. Precipitation was lower in both years but displayed a large variability.\\
We investigate the difference in uptake of ozone using the $\mathrm{DO_3SE}$ model, for typical vegetation at the location. We find that bespoke parameterization for subarctic vegetation is essential for an accurate estimation of Phyto-toxic Ozone Dose (POD) and an assessment of damage risk. Accurate estimates of the start and length of growing season are also crucial (up to $15\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ deviation in $\mathrm{POD_1}$).\\
We find that temperature and its derivative vapor pressure deficit are likely the most important climate variables while the variability of ozone concentration within climatological reasonable limits plays a minor role.
The use of subarctic parameterization suggests an improvement in risk assessment as POD values are increased compared to standard expectations; corroborated by the evidence of visible injury in the Svanhovd ozone garden. We estimate a total biomass reduction in Norway spruce of $2.5-3\,\unit{\%}$, downy birch by $15.5-17.4\,\unit{\%}$ and an above-ground biomass reduction in perennial grassland of $12.4-13.7\,\unit{\%}$.  
\end{abstract}

@@

Ground level ozone is a highly toxic pollutant known to cause damage to a variety of ecosystems around the World \citep{PT:Emberson2020}. At continuously high concentration levels, ozone negatively affects life expectancy of humans \citep{WHO2008} and is a cause of visible injury, photosynthetic damage, early senescence as well as programmed cell death of plants \citep{PCE:Kangasjarvi2005}. Annual global yield losses of four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} and a suggested loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
The crop yield loss studies indicate a threat to food security in rapidly developing countries, e.g., in East and South-East Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018}, where ozone concentrations related to enhanced pollutant emissions from the transport, residential, power generation and industrial sectors have increased in the past few decades. Even though long-term (pre 1950) observations of ground level ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase until relatively recently \citep[Chapter 2]{IPCC2013}. Recent trend analyses indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric ozone started to level-off or even decline in Europe \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}. This can be attributed to a successful implementation of air quality regulations coupled with economic restructuring which saw a move from coal to gas, reducing the number of episodes of peak concentrations especially in summertime over Europe and North America \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}. However, there are indications that an increase in the frequency of heatwave induced thermal stress will reduce the vegetations' capacity for ozone uptake and therefor diminish these achievements \citep{NCC:Lin2020}.\\

@@

For computing the $\mathrm{POD_y}$, an estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf. Details specific to the $\mathrm{DO_3SE}$ model are given in Section~\ref{sec:do3se}. $\mathrm{POD_y}$ is then calculated according to:
%
\begin{equation}
  \mathrm{POD_y} = \int{(\Phi_\mathrm{sto}-y)\cdot \mathrm{dt}},
  \label{eq:pod}
\end{equation}
%
with the hourly averaged ground level ozone flux through the stomata Eq.~(\ref{eq:flux_stomata}) and a flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}. The flux threshold $\mathrm{y}$ is representing the detoxification potential of the plant and is typically only exceeded during daylight hours (i.e. when global radiation is above $50\,\unit{W\,m^{-2}}$).

@@

Within the scope of this study to assess probable effects of conditions unique to northern Fennoscandia on ozone uptake by plants and resultant enhanced susceptibility to ozone inflicted damage, we conducted ground level ozone concentration measurements at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in the growing seasons of 2018/19 in collaboration with the Norwegian Institute for Air Research (NILU). 
In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden at Svanhovd. In contrast, no such damage was found on the clovers in 2019, although it was found on the sensitive tobacco cultivar. This indicates that the vegetation may have been more affected by ozone in 2018 than in 2019. In terms of $\mathrm{AOT_{40}}$, the threshold for protection of vegetation of $3000\,\unit{ppb}$ (integrated over the growing season) given by the EU directive has been well exceeded at many ozone monitoring sites in Southern Norway in the 2018 growing season \citep{NILU2019} but not at Svanhovd.

@@

The observation site at NIBIO Environment Centre Svanhovd and all related data are presented in Section~\ref{sec:data}. In Section~\ref{sec:stats}, we derive climatologies for ground level ozone concentration at Svanhovd and northern Fennoscandia based on in situ observations and also compute climatologies of near-surface temperature, precipitation, and global radiation for Svanhovd. Based on these, we evaluate anomalies of environmental key variables in 2018/19 by statistical means. In addition, the derived ground level ozone concentration climatology for northern Fennoscandia will be compared with available ozone reanalysis products. We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate main environmental drivers for the observed ozone damage in 2018 (Section~\ref{sec:do3se}). In this context, we introduce the$\mathrm{DO_3SE}$ model equations and derive bespoke parameterizations for subarctic species to assess systematic uncertainties arising from the use of default parameterizations. In Section~\ref{sec:conc} finally, we will summarize our results, draw conclusions, and give an outlook on further work.

@@

\subsection{Other data}
\label{subsec:other_data}
To determine the significance of the ozone conditions in 2018, we derive a ground level ozone concentration climatology for Svanvik. From 1986--1996 ozone monitoring has been operated by NILU at Svanhovd, but no ozone measurements have been conducted for the past $14\,\unit{years}$. Due to probable changes in northern hemispheric tropospheric background ozone and a reduction of summertime peak values related to ozone precursors which fell under air quality regulations in Europe, we will cross-calibrate our data with long-term ozone observations at other sites in northern Fennoscandia. In regard to ozone, available observational data are sparse in this region. An overview over the geographical locations of conducted atmospheric monitoring in the past and present is given in Fig.~\ref{fig:station_map_fennoscandia}. The data are described in Section~\ref{subsubsec:ebas}. 
As additional complement to observational data which are not available at all times and spatially sparse, we also evaluate various ozone reanalysis products (Section~\ref{subsubsec:ozone_rea}).

@@

\subsubsection{Ground level ozone observations}
\label{subsubsec:ebas}
All ground level ozone observations used in this study are available from the EBAS database operated by \citet{NILU_EBAS}. In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking. All stations with available long-term observations are located on higher ground than Svanhovd which will result in higher ground level ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors \citep[e.g.,][]{AB:Klingberg2009}. The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, and heath. However, at close vicinity it varies substantially. Ozone measurements at Pallas are taken on top of mount Sammaltunturi $100\,\unit{m}$ above tree line (pine forest) and mainly consists of low vascular plants, mosses, and lichen \citep{BER:Hatakka2003}. At Svanhovd, the ozone measurements are taken amidst a meadow surrounded by agriculturally managed herbaceous plants and grass (c.c. Fig.\ref{fig:svanhovd_research_station}).

@@

Details of reanalysis data sets referred to in this work are listed in Table~\ref{tab:ozone_rea}.
There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone (MACC and CAMSRA) \citep{ACP:Inness2013, ACP:Inness2019}. The temporal as well as spatial resolution of these reanalysis products is rather coarse: 3-hourly and $0.75^\circ\times 0.75^\circ$ or roughly $29.3\,\unit{km}\times 83.4\,\unit{km}$ at the location of Svanvik, respectively. In addition, higher resolution ground level ozone reanalysis for whole Europe is available from Copernicus service for regional air quality control (note: since May 31st 2020 through the CAMS Atmosphere Data Store (ADS)). The regional air quality model reanalysis ensemble (CAMS regional Ensemble) is based on nine European state-of-the-art numerical air quality models \citep{Copernicus_Regio_AIRQ}. The ensemble mean is at fairly high spatial and temporal resolution compared to the global reanalysis from ECMWF: $0.1^\circ\times 0.1^\circ$ or roughly $3.9\,\unit{km}\times 11.1\,\unit{km}$ and 1-hourly, respectively. The covered time periods differ and no data are available for before the turn of the millennium. The CAMSRA and CAMS Regional Air Quality ensemble mean are available in near real time, but only former covers a period of sufficient length for climate analysis (2003 -- present). All reanalyses use different but, at that time, latest versions of the operational weather forecast system (OpenIFS) from ECMWF. They differ substantially in the assimilated observational ozone data. The MACC reanalysis assimilates only satellite derived tropospheric column ozone (TC), while CAMSRA also includes ozone profiles from satellite retrievals. Most importantly for our study, in situ observations from ozone near-surface station networks are only assimilated in the CAMS Regional Ensemble. The quality with respect to observations in northern Fennoscandia will be discussed in Section~\ref{subsubsec:clim_ozone}. Without too much foreshadowing, the results prevented us from fetching the whole CAMSRA data.

@@

Finally, we shall look at the comparison of climatologies derived from observational ozone data and the various reanalysis products (Fig.~\ref{fig:ozone_climatologies}). As representative for the sites, the nearest neighboring grid point has been chosen. The ECMWF products do not reproduce observed ground level ozone well. The MACC reanalysis (Fig.~\ref{fig:ozone_climatologies}a) is in general too low ($(9\pm 7)\,\unit{ppb}$) and displays an erroneous seasonal cycle. Ozone concentrations peak already in March followed by a second peak in July. Also CAMSRA matches the observed ozone climatology poorly (Fig.~\ref{fig:ozone_climatologies}b). Despite displaying only one ozone peak, it still does not reproduce the seasonality in northern Fennoscandia well. The CAMSRA spring peak appears to lag behind by $1\,\unit{month}$ and is $5\,\unit{ppb}$ lower, whereas the minimum occurs in January compared to August/September. The annual amplitude ($(26\pm 1)\,\unit{ppb}$) is larger than in the climatology derived from observations ($19\,\unit{ppb}$). Modeled \chem{[O_3]} at Svanvik are even slightly higher than at Pallas. In contrast, the ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia quite well (Fig.~\ref{fig:ozone_climatologies}c), but is on average still slightly too low ($(2.8\pm 0.5)\,\unit{ppb}$).

\begin{figure}[t]
  %\includegraphics[width=8.3cm]{ozone_climatologies_portrait}
  \includegraphics[width=8.3cm]{fig07}
  \caption{Comparison of climatologies derived from observational data (EBAS) with various ozone reanalysis products: (a) ECMWF: MACC reanalysis; (b) ECMWF: CAMS reanalysis; (c) Copernicus: Regional model reanalysis ensemble mean. As representative for the sites, the nearest neighboring grid point has been chosen. The ECMWF products do not reproduce observed ground level ozone well. The MACC reanalysis is in general too low and displays an erroneous seasonal cycle. The CAMS reanalysis matches better but does not reproduce the seasonality in northern Fennoscandia. The ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia, though the modeled ozone concentrations are slightly too low.}
  \label{fig:ozone_climatologies}
\end{figure}

@@

In the following, we will use the interchangeable terms anomalies and residuals to distinguish between a climatological (anomalies) and a statistical (residuals) point of view.
All the above factors make it necessary to look not only at anomalies but statistically assess ozone concentration residuals at Svanvik as well as at Esrange and Pallas in comparison. Since the climatology of latter is more reliable, we will start with ground level ozone concentration residuals \chem{\Delta[O_3]} at Esrange and Pallas in 2018. As data for 2019 has not yet been available from EBAS (last access, February 2020), we refrain from looking into these.

@@

Because the available reanalysis products for ground level ozone have their limitations, we deem it necessary to reconstruct the missing data in 2018 for the purpose of $\mathrm{DO_3SE}$ modeling by applying a Reynolds's decomposition of the form:
\begin{equation}
  \chem{[O_3]} = \chem{\left<[O_3]\right>} + \chem{\Delta[O_3]},
  \label{eq:reynolds_decomp}
\end{equation}
with climatology \chem{\left<[O_3]\right>} and anomalies \chem{\Delta[O_3]}.
The exact procedure will be described in the following.
We derive hourly climatologies for individual stations and combine Esrange/Pallas to derive a climatology for northern Fennoscandia for July (Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}a). We compute temporal correlations between Svanvik and the other stations in northern Fennoscandia to estimate a time lag (Supplement Fig.~\ref{fig:time_lag_correlation}. We find $3\,\unit{h}$ for Esrange and Pallas with respect to Svanvik. Due to the high correlation with Svanvik (Appendix~Fig.~\ref{fig:density_distribution}), we decide on Pallas as reference station for the reconstruction. We correct the Fennoscandic climatology for the found time lag \chem{\left<[O_3]\right>_{hourly,\,t-3}} by shifting the respective time series back in time by $3\,\unit{h}$ and use it to derive a projection for Svanvik:
\begin{equation}
  X = \frac{\chem{\left<[O_3]\right>_{hourly}^{Svanvik}}}{\chem{\left<[O_3]\right>_{hourly,\,t-3}}}.
  \label{eq:mapping}
\end{equation}
As an intermediate step, we apply Eq.~(\ref{eq:reynolds_decomp}) to derive hourly anomalies for each station $i$
\begin{equation}
  \chem{\Delta[O_3]_{hourly}}^i = \chem{[O_3]_{hourly}}^i-\chem{\left<[O_3]\right>_{hourly}},
\end{equation}
with $i \in \{\text{Esrange, Pallas, Svanvik\}}$.
In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}b), these hourly ozone concentration anomalies are shown. We also correct the ozone anomalies for Pallas for the time lag \chem{\Delta[O_3]_{hourly,\,t-3}^{Pallas}} and use the mapping function in Eq.~(\ref{eq:mapping}) to derive reconstructed anomalies for the missing values at Svanvik:
\begin{equation}
    \chem{\Delta [O_3]_{hourly}^{Svanvik,\,reco}} = \chem{\Delta[O_3]_{hourly,\,t-3}^{Pallas}} \cdot X
\end{equation}
Finally, we add the bias corrected Svanvik ozone climatology and derive a reconstructed ozone concentration time series \chem{[O_3]_{hourly}^{Svanvik,\,reco}}. In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}c), the result is shown together with the observed data and the Copernicus Regional Model Reanalysis at the nearest grid point for comparison.

@@

We found representative ozone climatologies (multi-annual means) for northern Fennoscandia and Svanhovd based on ground level ozone monitoring. To account for missing data from our 2018 record, we proposed a routine for gap filling of ozone observations following a Reynolds's decomposition into anomalies and climatology. We also assessed the quality of available global reanalysis (ECMWF MACC and CAM) and regional model reanalysis (Copernicus) ground level ozone concentrations products and conclude that only the Copernicus product can sufficiently reproduce observed ozone concentrations in northern Fennoscandia.
